---
title: "week5_exercise"
author: "Wenke Zimmermann"
date: "2022-05-21"
output: html_document
---

## Libraries ###
```{r message=FALSE, warning= FALSE}
library(readr)
library(udunits2)
library(sf)
library(terra)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tmap)

```

## Load data as an sf-object
```{r message=FALSE, warning= FALSE}

wildschwein <- read_delim("wildschwein_BE_2056.csv",",") %>%
  st_as_sf(coords = c("E", "N"), crs = 2056, remove = FALSE)

```


## Tasks 1: Import and visualize spatial data
```{r message=FALSE, warning= FALSE}

feldaufnahmen <- read_sf("Feldaufnahmen_Fanel.gpkg")

#What information does the dataset contain?
feldaufnahmen
#Geometrie: Polygon
#Projektion: Landeskoordinaten
#975 Zeilen, 3 Attribute
#...

#What is the geometry type of the dataset (possible types are: Point, Lines and Polygons)?
#Polygons

#What are the data types of the other columns?
#FieldID: double (Nummer des Feldes)
#geom: Koordinaten

#What is the coordinate system of the dataset?
#Landeskoordinaten

```


## Task 2: Annotate Trajectories from vector data
what crop was visited by which wild boar, and at what time
filter wildboar data to may to june (crop data is most relevant in summer), save output to new variable
```{r message=FALSE, warning= FALSE}

#Anlegen convenience-Variable 'month'
wildschwein$month <- month(wildschwein$DatetimeUTC) #einfache Variante um convenience-Variable anzulegen

#alternativ mit mutate
#wildschwein %>%
#  mutate(month ???? )


#Filter auf Monate Mai und Juni
wildschwein_sommer <- wildschwein %>%
  filter(month == 05 | month == 06)

#folgende Variante ist viiiiieeeel zu umständlich
#wildschwein_sommer2015 <- wildschwein %>%
#  filter(DatetimeUTC >= as.POSIXct("2015-05-01", tz = "UTC") & DatetimeUTC <= #as.POSIXct("2015-06-30", tz = "UTC"))

#wildschein_summer2014 <- wildschwein %>%
#    filter(DatetimeUTC >= as.POSIXct("2014-05-01 ", tz = "UTC") & DatetimeUTC <= #as.POSIXct("2014-06-30", tz = "UTC"))



#Overlay the filtered dataset with your fanel data to verify the spatial overlap
#To sematically annotate wildboar locations with crop information, use a spatial join

wildschwein_overlay <- st_join(wildschwein_sommer, feldaufnahmen) 

```


## Task 3: Explore annotated trajectories
Think of ways you could visually explore the spatio-temporal patterns of wild boar in relation to the crops. In our example below we visualize the percentage of samples in a given crop per hour.
```{r message=FALSE, warning= FALSE}

wildschwein_overlay$time <- 
  hour(wildschwein_overlay$DatetimeUTC)

#alternativ mit mutate
#wildschwein_overlay %>%
#  mutate(timeRound = round_date(DatetimeUTC, unit = "1 hour"))


wildschwein_sry <- wildschwein_overlay %>%
  group_by(TierName, Frucht) %>%
  summarise(Summe = sum(time))



ggplot()+
  geom_bar(wildschwein_sry) +
  facet_wrap(~TierName)

 

y = (value)/sum(value)
y = Frucht/sum(Frucht)

plot_data = mdat %>% group_by(group, variable) %>% summarise(total = sum(value))

ggplot(wildschwein_sry, aes(fill = Frucht)) +  
  geom_bar(aes(x = time, 
               y = (Frucht)/sum(Frucht)), 
               stat = "identity", 
               position = "dodge") + 
  scale_y_continuous(labels = scales::percent) + 
  xlab("Time (Rounded to neares hour") + 
  ylab("Percentage")



```



## Task 4: Import and visualize vegetationindex (raster data)
Import this dataset vegetationshoehe_LFI.tif
It's a prepared Vegetation Height Model provided by the Swiss National Forest Inventory (NFI) and contains high resolution information (1x1 Meter) on the vegetation height, which is determined from the difference between the digital surface models DSM and the digital terrain model by swisstopo (swissAlti3D). Buildings are eliminated using a combination of the ground areas of the swisstopo topographic landscape model (TLM) and spectral information from the stereo aerial photos.
```{r message=FALSE, warning= FALSE}

#Import tif-data
vegetation <- terra::rast("vegetationshoehe_LFI.tif")


#Visualize the raster data using tmap (ggplot is too slow for raster data)
plot(vegetation)


## Task 5: Adding a background map (library tmap)
tm_shape(vegetation)+
  tm_raster()
#Raster wurde autoamtische gedownsampelt
#Farbe ist gelb anstatt blau

```


