---
title: "week5_exercise"
author: "Wenke Zimmermann"
date: "2022-05-21"
output: html_document
---

## Libraries ###
```{r message=FALSE, warning= FALSE}

library(readr)
library(udunits2)
library(sf)
library(terra)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tmap)

```

## Load data as an sf-object
```{r message=FALSE, warning= FALSE}

wildschwein <- read_delim("wildschwein_BE_2056.csv",",") %>%
  st_as_sf(coords = c("E", "N"), crs = 2056, remove = FALSE)

```


## Tasks 1: Import and visualize spatial data
```{r message=FALSE, warning= FALSE}

feldaufnahmen <- read_sf("Feldaufnahmen_Fanel.gpkg")

#What information does the dataset contain?
feldaufnahmen
#Geometrie: Polygon
#Projektion: Landeskoordinaten
#975 Zeilen, 3 Attribute
#...

#What is the geometry type of the dataset (possible types are: Point, Lines and Polygons)?
#Polygons

#What are the data types of the other columns?
#FieldID: double (Nummer des Feldes)
#geom: Koordinaten

#What is the coordinate system of the dataset?
#Landeskoordinaten

```


## Task 2: Annotate Trajectories from vector data
what crop was visited by which wild boar, and at what time; filter wildboar data to may to june (crop data is most relevant in summer), save output to new variable
```{r message=FALSE, warning= FALSE}

#Anlegen einer convenience-Variable, die die Monate enthält
wildschwein  <- mutate(wildschwein, month = month(DatetimeUTC))

#alternativ
#wildschwein$month <- month(wildschwein$DatetimeUTC) #sehr einfache Variante um convenience-Variable anzulegen


#Filter auf Monate Mai und Juni
wildschwein_filter <- wildschwein %>%
  filter(month == 05 | month == 06)

#folgende Variante ist viiiiieeeel zu umständlich
#wildschwein_filter <- wildschwein %>%
#  filter(DatetimeUTC >= as.POSIXct("2015-05-01", tz = "UTC") & DatetimeUTC <= #as.POSIXct("2015-06-30", tz = "UTC")) |
#  filter(DatetimeUTC >= as.POSIXct("2014-05-01 ", tz = "UTC") & DatetimeUTC <= #as.POSIXct("2014-06-30", tz = "UTC"))


#Overlay the filtered dataset with your fanel data to verify the spatial overlap
#To sematically annotate wildboar locations with crop information, use a spatial join

wildschwein_crop_overlay <- st_join(wildschwein_filter, feldaufnahmen)

```


## Task 3: Explore annotated trajectories
Think of ways you could visually explore the spatio-temporal patterns of wild boar in relation to the crops.
```{r message=FALSE, warning= FALSE}

#Versuch zur Darstellung, der nicht bis zum Ende funktionierte

#Zähle Häufigkeit je Anzahl crop-Typ
#wildschwein_crop_overlay %>%
#  add_count(Frucht)

#wildschwein_crop_overlay %>%
#  mutate(wildschwein_crop_overlay, AnzFrucht = count(Frucht))

#wildschwein_crop_overlay %>%
#  group_by(Frucht) %>%
#  summarise(AnzFruch = n())

#neue Variable mit gerundeten Stunden
#wildschwein_crop_overlay <- wildschwein_crop_overlay %>%
#  mutate(time = round_date(DatetimeUTC, unit = "1 hour"))

#Zähle Häufigkeit je Anzahl crop-Typ
#wildschwein_crop_sum <- wildschwein_crop_overlay %>%
#  group_by(Frucht) %>%
#  summarise(AnzFruch = n())
  
#Visualisiere Option 1 (das mit der Frucht funzt noch nicht!!!)
#ggplot(wildschwein_crop_overlay, aes(x = time, y = Frucht)) +
#  geom_bar() + 
#  scale_y_continuous(labels = scales::percent) +
#  xlab("Time (Rounded to the nearest hour)") + 
#  ylab("Percentage") +
#  facet_wrap(~TierName)



# Alternative von Nils
library(forcats)  #zum Handling von Factors

wildschwein_smry <- wildschwein_crop_overlay %>%
  st_set_geometry(NULL) %>%
  mutate(
    hour = hour(round_date(DatetimeUTC,"hour")),
    Frucht = ifelse(is.na(Frucht),"other",Frucht),
    Frucht = fct_lump(Frucht, 5, other_level = "other"),
  ) %>%
  group_by(TierName, hour, Frucht) %>%
  count() %>%
  ungroup() %>%
  group_by(TierName , hour) %>%
  mutate(perc = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Frucht = fct_reorder(Frucht, n,sum, desc = TRUE)
  )


p1 <- ggplot(wildschwein_smry, aes(hour,perc, fill = Frucht)) +
  geom_col(width = 1) +
  scale_y_continuous(name = "Percentage", labels = scales::percent_format()) +
  scale_x_continuous(name = "Time (rounded to the nearest hour)") +
  facet_wrap(~TierName ) +
  theme_light() +
  labs(title = "Percentages of samples in a given crop per hour",subtitle = "Only showing the most common categories")

p1

p1 +
  coord_polar()  +
  labs(caption = "Same visualization as above, displayed in a polar plot")


```



## Task 4: Import and visualize vegetationindex (raster data)
The dataset vegetationshoehe_LFI.tif is a prepared Vegetation Height Model provided by the Swiss National Forest Inventory (NFI) and contains high resolution information (1x1 Meter) on the vegetation height, which is determined from the difference between the digital surface models DSM and the digital terrain model by swisstopo (swissAlti3D). Buildings are eliminated using a combination of the ground areas of the swisstopo topographic landscape model (TLM) and spectral information from the stereo aerial photos.
```{r message=FALSE, warning= FALSE}

#Import tif-data
vegetation <- terra::rast("vegetationshoehe_LFI.tif")


#Visualize the raster data using tmap (ggplot is too slow for raster data)
plot(vegetation)


## Task 5: Adding a background map (library tmap)
tm_shape(vegetation)+
  tm_raster()
#Raster wurde automatisch gedownsampelt
#Farbe ist gelb anstatt blau


# Alternativ von Nils
tm_shape(vegetation) + 
  tm_raster(palette = "viridis",style = "cont", legend.is.portrait = FALSE) +
  tm_layout(legend.outside = TRUE,legend.outside.position = "bottom", frame = FALSE)

```


## Task 5: Annotate Trajectories from raster data
Semantically annotate your wildboar locations with the vegetation index.
```{r message=FALSE, warning= FALSE}

#Since you are annotating a vector dataset with information from a raster dataset, you cannot use st_join but need the function extract from the terra package

#Variante von Nils
wildschwein_vegetation_overlay <- terra::extract(vegetation, st_coordinates(wildschwein))

wildschwein <- cbind(wildschwein, wildschwein_vegetation_overlay)

wildschwein


```
